---
title: "Intervention_Analysis_2016_Rio_Olympics"
author: "Prafulla Ranjan Dash"
output: html_document
---

#### Here, we try and apply Intervention Analysis for the number of views time-series for the *2016_Summer_Olympics_en* webpage.

```{r workdirimp, echo=TRUE, message=FALSE, warning=FALSE}
# basic imports
library("fpp", lib.loc="C:/Users/HP/Anaconda3/envs/rstudio/lib/R/library")
library("ggplot2", lib.loc="C:/Users/HP/Anaconda3/envs/rstudio/lib/R/library")
library("stats", lib.loc="C:/Users/HP/Anaconda3/envs/rstudio/lib/R/library")
library("zoo", lib.loc="C:/Users/HP/Anaconda3/envs/rstudio/lib/R/library")
library("TSA", lib.loc="C:/Users/HP/Anaconda3/envs/rstudio/lib/R/library")
```

#### Importing the data

```{r dataimport, echo=TRUE, message=FALSE, warning=FALSE}
setwd("D:/USMS/UChicago/STUDIES/Autumn 2020/MSCA 31006 1 Time Series Analysis and Forecasting/Final Project/Data")
page_2016_Rio_Olympics_views <- data.matrix(read.csv("train_1.csv", header = TRUE, row.names = 1,sep = ",",nrows = 1,skip = 12186))
```

#### Preparing data to make it ready for time-series

```{r dataprepare, echo=TRUE, message=FALSE, warning=FALSE}
# removing column names
dimnames(page_2016_Rio_Olympics_views) <- NULL

# converting to array
page_2016_Rio_Olympics_views <- array(page_2016_Rio_Olympics_views)

# checking the values
head(page_2016_Rio_Olympics_views)

# no. of ovservations
length(page_2016_Rio_Olympics_views)
```

Looking at the entire data - 

```{r view, echo=TRUE, message=FALSE, warning=FALSE}
plot(page_2016_Rio_Olympics_views,type = 'l')
```

Looking at this dataset, we can see there are few time periods where there seems to be a clear intervention at t=175. So, we will take a sub-section of this data for our analysis.

```{r split, echo=TRUE, message=FALSE, warning=FALSE}
page_2016_Rio_Olympics_views <- page_2016_Rio_Olympics_views[100:400]

# converting data to ts
page_2016_Rio_Olympics_views <- ts(page_2016_Rio_Olympics_views,frequency = 1)

# plotting the time-series
autoplot(page_2016_Rio_Olympics_views)
```

So, here we can clearly see some sort of **intervention** that took place at around the time-period of 75 days for this chunk of data (175 days overall). There is a clear **upward trend** after the time-period of 75 days. 

Now, since the intervention is at time-period 75, we will divide our data into data-points prior to time-period 75 and post time-period 75. We will do this so that we can find the **underlying process** that is present in our observations, which are a **mix of the underlying process and the intervention**.

Also, we will split our data to train and test so that we can see how our model performs on new data.

```{r split2, echo=TRUE, message=FALSE, warning=FALSE}
# train-test split
#page_2016_FIFA_Club_World_Cup_views_train <- page_2016_FIFA_Club_World_Cup_views[1:340]
#page_2016_FIFA_Club_World_Cup_views_test <- page_2016_FIFA_Club_World_Cup_views[341:350]

# splitting to prior and post intervention
page_2016_Rio_Olympics_views_prior <- page_2016_Rio_Olympics_views[1:75]
page_2016_Rio_Olympics_views_post <- page_2016_Rio_Olympics_views[76:300]

# converting to time-series
page_2016_Rio_Olympics_views_prior <- ts(page_2016_Rio_Olympics_views_prior)
page_2016_Rio_Olympics_views_post <- ts(page_2016_Rio_Olympics_views_post)
```

Now, let's see how our prior intervention data looks

```{r prior, echo=TRUE, message=FALSE, warning=FALSE}
autoplot(page_2016_Rio_Olympics_views_prior)
```


Looking at the data - 

1. There is no visible non-seasonal trend.  
2. There might be some seasonality where some patterns are repeating themselves.
3. The variance seems unstable, might require some Box-Cox transformation.

Applying Box-Cox - 

```{r boxcox, echo=TRUE, message=FALSE, warning=FALSE}
BoxCox.lambda(page_2016_Rio_Olympics_views_prior)

page_2016_Rio_Olympics_views_prior_transformed <- BoxCox(page_2016_Rio_Olympics_views_prior,lambda = -0.6576963)

autoplot(page_2016_Rio_Olympics_views_prior_transformed)
```

The variance now seems to have stabilized a bit. Now, let's look at the ACF and PACF for this signal.

```{r acf, echo=TRUE, message=FALSE, warning=FALSE}
tsdisplay(page_2016_Rio_Olympics_views_prior_transformed)
```

We can see from the ACF that our data is **not stationary** as there are significant spikes along with a sinusoidal component. Let's apply one order of non-seasonal differencing.

```{r diff, echo=TRUE, message=FALSE, warning=FALSE}
# differencing
page_2016_Rio_Olympics_views_prior_transformed_diff <- diff(page_2016_Rio_Olympics_views_prior_transformed)

# visulalizing
tsdisplay(page_2016_Rio_Olympics_views_prior_transformed_diff)
```

one order of seasonal differencing


```{r sdiff, echo=TRUE, message=FALSE, warning=FALSE}
# differencing
page_2016_Rio_Olympics_views_prior_transformed_diff2 <- diff(page_2016_Rio_Olympics_views_prior_transformed_diff,lag = 7)

# visulalizing
tsdisplay(page_2016_Rio_Olympics_views_prior_transformed_diff2)
```


We see that the data still does not appear stationary.

Let's allow auto.arima to decide the best model for this data.

```{r m_prior, echo=TRUE, message=FALSE, warning=FALSE}
m_prior <- auto.arima(page_2016_Rio_Olympics_views_prior,lambda = 'auto')
checkresiduals(m_prior)
```

```{r model, echo=TRUE, message=FALSE, warning=FALSE}
m_prior_forecast <- forecast(m_prior,h = 225)
plot(m_prior_forecast)
```

```{r model11, echo=TRUE, message=FALSE, warning=FALSE}
#plot(stepfun(1,1:2))
exp_intervention=1/dexp(1:226,rate = 0.04)
plot(exp_intervention)
m1 <- arimax(log(page_2016_Rio_Olympics_views),order = c(5,1,2),
             xreg = data.frame(x=c(rep(0,75),exp_intervention)),method = 'ML')
m1
m2 <- arimax(log(page_2016_Rio_Olympics_views),order = c(5,1,2),
              xtransf = data.frame(x=c(rep(0,75),exp_intervention)), method = 'ML')
m3 <- arimax(page_2016_Rio_Olympics_views,order = c(5,1,2),
             xreg = data.frame(x=c(rep(0,75),1:226)),method = 'ML')

m4 <- arimax(page_2016_Rio_Olympics_views,order = c(5,1,2),
             xreg = data.frame(x=c(rep(0,75),1:226)),method = 'ML')

checkresiduals(m2)
```

```{r diag, echo=TRUE, message=FALSE, warning=FALSE}
checkresiduals(m1)
plot(1/dexp(1:226,rate = 0.03))
```
