---
title: "Web Traffic Time Series Forecasting -"
author: "Surendiran Rangaraj"
date: "11/7/2020"
output:
  html_document:
    fig_height: 5
    fig_width: 8
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: false
    number_sections: false
    theme: readable
    highlight: tango
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(fpp2) # time series analysis
library(tseries) # time series analysis
library(ggplot2) # visualization
library(forecast)  # time series analysis
library(TSA)  # time series analysis
library(dplyr) # data manipulation
library(readxl) # data input
library(xts) # time series analysis
library(tibble) # data wrangling
library(tidyr) # data wrangling
library(prophet) # time series analysis
library(zoo)
library(TSstudio)
library(lubridate)

```


```{r Read data}
#Read the Input data as Dataframe
webdata_path = 'C:/Users/rsure/Desktop/Uchicago/Time Series/Final Project/'
#train_1_df = read.table(paste0(webdata_path,'train_1.csv'),sep=',',header = TRUE,skip =0,row.names = "Page")
#train_2_df = read.table(paste0(webdata_path,'train_2.csv'),sep=',',header = TRUE,skip =0,row.names = "Page")
olympics_web_df = read.table(paste0(webdata_path,'2016_Olympics_train1.csv'),sep=',',header = TRUE,skip =0)
#
c(ncol(olympics_web_df),nrow(olympics_web_df))
```

```{r data exploration}
print(paste0("Number of Missing Values for 2016 Olympics:",(sum(is.na(olympics_web_df))/(ncol(olympics_web_df)*nrow(olympics_web_df)))))
olympics_web_array=array(olympics_web_df[,'Visits'])
length(olympics_web_array)
head(olympics_web_array)

#Plot

plot_olympics <- data.frame(myts = olympics_web_array,
                  Date = seq(as.Date("2015-07-01"), as.Date("2016-12-31") , by = 1))
plot(olympics_web_array,type = 'l')
plot(plot_olympics[c('Date','myts')][100:135,],type = 'l')



# convert to time-series
idx <- seq(as.Date("2015-07-01"), as.Date("2016-12-31"), by = "day")
olympics_web_ts = ts(olympics_web_array, frequency = 7)
olympics_wed_ts = ts(olympics_web_df[,c('Visits','Weekday','Period')], frequency = 7)
head(olympics_web_ts)
head(olympics_web_ts)
tsdisplay(olympics_web_ts,main = 'Time Series Display of 2016 Olympics' )

#Seasonal Decomposition
#olympics_web_ts[,'Visits'] %>% mstl() %>% autoplot() 
olympics_web_df[,'Visits'] %>% mstl() %>% autoplot() 
olympics_wed_ts[,'Visits'] %>% mstl(s.window = "periodic",robust=TRUE) %>% autoplot() 

plot(seasadj(mstl(olympics_wed_ts[,'Visits'],s.window = 7,robust=TRUE)))

plot(decompose(olympics_wed_ts[,'Visits'],type = 'additive'))
plot(decompose(olympics_wed_ts[,'Visits'],type = 'multiplicative'))

ggseasonplot(olympics_wed_ts[,'Visits'], year.labels=TRUE, year.labels.left=TRUE) +
  ylab("Visits") +
  ggtitle("Seasonal plot: Olympics visits")

```
No Missing values found of the 550 Observations . 

### 2016 Summer Olympics Time series modeling 

There is block intervention in our data. We will look at the time series data for 2016 Summer Olympics and perform basic modeling

### Convert to Time Series
First Let's convert it to time-series and then split it to train and test - we will train our models for the period 1st July 2015 till 30th Nov 2016 and we will forecast for 1st Dec 2016 to 31st Dec 2016


```{r split}

split_ts <- ts_split(ts.obj = olympics_web_ts, sample.out = 31)

olympics_web_ts_train <- split_ts$train
olympics_web_ts_test <- split_ts$test

# viewing the train and test
tail(olympics_web_ts_train) 
head(olympics_web_ts_test)

print(paste0("Total Observations in Training dataset : ", length(olympics_web_ts_train), " Total Observations in Test dataset : ", length(olympics_web_ts_test)))

autoplot(olympics_web_ts_train,xlab = "Time",ylab = "No of Visits", main = "Time Series plot of 2016 Olympics English Web page Visits")
autoplot(olympics_web_ts_test,xlab = "Time",ylab = "No of Visits", main = "Time Series plot of 2016 Olympics English Web page Visits")

#time series display
tsdisplay(olympics_web_ts_train,ylab="2016 Rio Summer Olympics Web page Visits-Train",xlab="Time")
tsdisplay(olympics_web_ts_test,ylab="2016 Rio Summer Olympics Web page Visits-Test",xlab="Time")

```

```{r test}
kpss.test(olympics_web_ts_train,null = 'Trend') # p-value = 0.01261  < 0.05 # Reject Null Hypothesis => The process is not stationary 
adf.test(olympics_web_ts_train) # p-value = 0.07705 > 0.05  # Reject Null Hypothesis => The process is not stationary 

print(paste0("Boxcox lambda value for the entire dataset  is :" ,BoxCox.lambda(olympics_web_ts_train)))
print(paste0("Boxcox lambda value for the train dataset  is :" ,BoxCox.lambda(olympics_web_ts_train)))

```

<div class="alert alert-info">
 
 
</div>


**Kwiatkowski Phillips Schmidt Shin (KPSS) tests**

Null hypothesis: The process is stationary => Large p value

Alternative hypothesis: The process is nonstationary =>  Small p value

**Augmented Dickey Fuller test**

Null hypothesis: The process is nonstationary => Large p value

Alternative hypothesis: The process is stationary => Small p value

### Remove Outliers

```{r outlier, include= FALSE}
```

### Box-Cox Transformation

```{r box-cox}

lambda = BoxCox.lambda(olympics_web_ts_train)
olympics_web_ts_Boxcox = BoxCox(olympics_web_ts_train,lambda)

tsdisplay(olympics_web_ts_Boxcox)
plot(olympics_web_ts_Boxcox[1:30],type="l")

kpss.test(olympics_web_ts_Boxcox) # p-value = 0.01  < 0.05 # Reject Null Hypothesis => The process is not stationary 
adf.test(olympics_web_ts_Boxcox) # p-value = 0.88  > 0.05 # Accept Null Hypothesis => The process is not stationary 

```
###  Differencing

```{r nsdiff,include = TRUE}

# Apply seasonal differencing
olympics_web_ts_Boxcox_sdiff <- diff(olympics_web_ts_Boxcox,7)
kpss.test(olympics_web_ts_Boxcox_sdiff)
tsdisplay(olympics_web_ts_Boxcox_sdiff,main = 'Time Series Display of Olympics Data- BoxCox Seasonal Differencing')

# Apply non-seasonal differencing
olympics_web_ts_Boxcox_diff <- diff(olympics_web_ts_Boxcox_sdiff)
tsdisplay(olympics_web_ts_Boxcox_diff,main = 'Time Series Display of Olympics Differenced Data')
kpss.test(olympics_web_ts_Boxcox_diff) # Null hypothesis: The process is stationary => Large p value of 0.1

# No seasonal - only non-seasonal differencing
olympics_web_ts_Boxcox_nsdiff <- diff(olympics_web_ts_Boxcox)
kpss.test(olympics_web_ts_Boxcox_nsdiff)
tsdisplay(olympics_web_ts_Boxcox_sdiff,main = 'Time Series Display of Olympics Data- Differencing')

# No seasonal - only non-seasonal differencing
olympics_web_ts_nsdiff <- diff(olympics_web_ts)
kpss.test(olympics_web_ts_Boxcox_nsdiff)
tsdisplay(olympics_web_ts_Boxcox_sdiff,main = 'Time Series Display NSeasonal Differencing-No Boxcox')

plot_olympics <- data.frame(myts = olympics_web_array,
                  Date = seq(as.Date("2015-07-01"), as.Date("2016-12-31") , by = 1))
plot(olympics_web_ts_Boxcox_nsdiff[1:35],type = 'l')

```

### ARFIMA 

```{r arfima}

acf(olympics_web_ts_Boxcox,lag.max = 150)
olympics_web_arfima = arfima(olympics_web_ts_Boxcox)
summary(olympics_web_arfima)

```


### ETS model

 ETS model and see how it turns out.

```{r ets}
olympics_web_ets_fit <- ets(olympics_web_ts_train,lambda = "auto")
olympics_web_ets_fit 

plot(forecast(olympics_web_ets_fit,h=10))
checkresiduals(olympics_web_ets_fit)

 olympics_web_ets_fc= forecast(olympics_web_ets_fit,h=31)

 olympics_web_ets_fc %>% autoplot(xlab = 'Time',ylab='Visits',main = '2016 Olympics Visits - ETS')
##
accuracy(olympics_web_ets_fc,olympics_web_ts_test)

```


### Naive method

```{r naive}

checkresiduals(naive(olympics_web_ts))
mean(naive(olympics_web_ts)$residuals,na.rm = TRUE)

```

### Spectral Analysis


```{r periodogram}
spectrum(olympics_web_df[,'Visits'])
olymp_pgrm = periodogram(olympics_web_df[,'Visits'],log="yes") # spike in low frequency means it is dictated by trend and no other frequency
round(olymp_pgrm$spec,2)
```

### Applying auto.arima
#### auto arima with no seasonal differencing 

```{r autoarima1}

olympics_web_arima_fit_ns <- auto.arima(olympics_web_ts_train,lambda = "auto",seasonal = FALSE)
olympics_web_arima_fit_ns
checkresiduals(olympics_web_arima_fit_ns)
```


### ARIMA with Seasonalality

```{r wd autoarima1}

olympics_web_arima_fit_d1 <- auto.arima(olympics_web_ts_train,lambda = "auto",seasonal = TRUE,d=1)
olympics_web_arima_fit_d1

plot(olympics_web_arima_fit_d1$x, col="red", type="l")
lines(fitted(olympics_web_arima_fit_d1), type="l",col="blue")
checkresiduals(olympics_web_arima_fit_d1)# p-value = 0.0191 < 0.05 ==> Reject Null hypothesis ==>  The TS is not independently distributed
Box.test(olympics_web_arima_fit_d1$residuals,type = c("Ljung-Box"),lag=20)
Acf(olympics_web_arima_fit_d1$residuals,lag.max = 60)

#forecast
olympics_df_arima_forecast_d1 = forecast(olympics_web_arima_fit_d1,h=31)
autoplot(olympics_df_arima_forecast_d1,main='Forecast with Seasonal Differencing')
accuracy(olympics_df_arima_forecast_d1,olympics_web_ts_test)


```

```{r wd  fitted1}

olympics_web_arima_fit_sd <- auto.arima(olympics_web_ts_train,lambda = "auto",seasonal = TRUE,d=1,D=1)
olympics_web_arima_fit_sd

plot(olympics_web_arima_fit_sd$x, col="red", type="l")
lines(fitted(olympics_web_arima_fit_sd), type="l",col="blue")
checkresiduals(olympics_web_arima_fit_sd)# p-value = 0.0191 < 0.05 ==> Reject Null hypothesis ==>  The TS is not independently distributed
Box.test(olympics_web_arima_fit_sd$residuals,type = c("Ljung-Box"),lag=20)
Acf(olympics_web_arima_fit_sd$residuals,lag.max = 60)

#forecast
olympics_df_arima_forecast_sd = forecast(olympics_web_arima_fit_sd,h=31)
autoplot(olympics_df_arima_forecast_sd,main='Forecast with Seasonal Differencing')
accuracy(olympics_df_arima_forecast_sd,olympics_web_ts_test)

```



```{r fourier ns non wd}

fr = fourier(olympics_web_ts_train,K = 3)
bestfit = list(aicc=Inf)

for (i in 1:3)
{
  olympics_df_arima_fit_frns <- auto.arima(olympics_web_ts_train,xreg=fourier(olympics_web_ts_train, K=i),lambda = "auto",seasonal = FALSE)
  if (olympics_df_arima_fit_frns$aicc < bestfit$aicc)
    bestfit = olympics_df_arima_fit_frns
  else break;
}

bestfit
#check residuals
checkresiduals(bestfit)
Box.test(bestfit$residuals,type = c("Ljung-Box"),lag=20)
Acf(bestfit$residuals,lag.max = 60)
summary(bestfit)

fc_fr = fourier(olympics_web_ts_train, K=3,h=31)
olympics_df_arima_forecast_frns = forecast(bestfit,xreg=fc_fr)
autoplot(olympics_df_arima_forecast_frns)
accuracy(olympics_df_arima_forecast_frns,olympics_web_ts_test)
#Forecast Plot
plot(olympics_web_arima_fit_sd$x, col="red", type="l")
lines(fitted(bestfit), type="l",col="blue")
```

rule-of-thumb could be modified a little as follows:

For non-seasonal time series, use  h = min (10,T/5).
For seasonal time series, use h=min(2m,T/5).


``` {r tbats fourier}

#olympics_tbats_fit = tbats(olympics_web_df[,'Visits'])

olympics_tbats_fit = tbats(olympics_web_ts_train)
olympics_tbats_fit

checkresiduals(olympics_tbats_fit)

olympics_tbats_fc = forecast(olympics_tbats_fit,h=31) 
accuracy(olympics_tbats_fc,olympics_web_ts_test)


#no short term ARMA model needed here


```

```{r cbind fc}

model_values = cbind(train_fitted=fitted(olympics_df_arima_forecast_frns),Dec2016_forecast = olympics_df_arima_forecast_frns$mean)
write.csv(model_values,'ARIMA_with_Errors_fitted_forecasted_values.csv')
```

###CV - Expanding and Sliding Window

#### Set/Initialize Parameters for the model
```{r set values}
#Set the CV parameters

k <- 160 # minimum data length for fitting a model
n <- length(olympics_web_ts) # Number of data points
p <- 7 ## Period
H <- 31 # Forecast Horizon

```
#### ARIMA and ETS Models

```{r CV arima,include=FALSE,eval=FALSE}
defaultW <- getOption("warn") 
options(warn = -1)

st <- tsp(olympics_web_ts_train)[1]+(k-2)/p #  gives the start time in time units, #1998.5

arima_mae_1 <- matrix(NA,n-k,H)
arima_mae_2 <- matrix(NA,n-k,H)
ets_mae_1 <- matrix(NA,n-k,H)
ets_mae_2 <- matrix(NA,n-k,H)

arima_rmse_1 <- matrix(NA,n-k,H)
arima_rmse_2 <- matrix(NA,n-k,H)
ets_rmse_1 <- matrix(NA,n-k,H)
ets_rmse_2 <- matrix(NA,n-k,H)

arima_aicc_1 <- matrix(NA,n-k)
arima_aicc_2 <- matrix(NA,n-k)
ets_aicc_1 <- matrix(NA,n-k)
ets_aicc_2 <- matrix(NA,n-k)


for(i in 1:(n-k))
#for(i in 1:25)
{
### One Month rolling forecasting
# Expanding Window 
expand_train1 <- window(olympics_web_ts, end=st + i/p)  ## Window Length: k+i
  
  # Sliding Window - keep the training window of fixed length. 
  # The training set always consists of k observations.
slide_train2 <- window(olympics_web_ts, start=st+(i-k+1)/p, end=st+i/p) ## Window Length: k
  
test <- window(olympics_web_ts, start=st + (i+1)/p, end=st + (i+H)/p) ## Window Length: H

 if (i==1 || i==80) {
  cat(c("*** CV", i,":","len(Expanding Window):",length(expand_train1), "len(Sliding Window):",length(slide_train2), "len(Test):",length(test),'\n'  ))
  cat(c("*** TRAIN -  Expanding WIndow:",tsp(expand_train1)[1],'-',tsp(expand_train1)[2],'\n'))
  cat(c("*** TRAIN - Sliding WIndow:",tsp(slide_train2)[1],'-',tsp(slide_train2)[2],'\n'))
  cat(c("*** TEST:",tsp(test)[1],'-',tsp(test)[2],'\n'))
  cat("*************************** \n \n")
}
  
  ##Seasonal ARIMA model - sARIMA([2,1,2][2,1,0]7 with Expanding Window
  arima_fit_1 <- Arima(expand_train1, order=c(1,1,4), xreg=fourier(expand_train1, K=3), lambda="auto", method="ML")
  fc_frex = fourier(expand_train1, K=3,h=H)
  arima_fcast_1 <- forecast(arima_fit_1, xreg=fc_frex)
  
  
  
  ##Seasonal ARIMA model - sARIMA([1,0,1][0,1,2]12 with Sliding Window
  arima_fit_2 <- Arima(slide_train2, order=c(1,1,4), xreg=fourier(slide_train2, K=3), lambda="auto", method="ML")
  fc_frsl = fourier(slide_train2, K=3,h=H)
  arima_fcast_2 <- forecast(arima_fit_2, xreg=fc_frsl)
  

# Calculate Forcast MAE
  arima_mae_1[i,1:length(test)] <- abs(arima_fcast_1[['mean']]-test)
  arima_mae_2[i,1:length(test)] <- abs(arima_fcast_2[['mean']]-test)

  
# RMSE sqrt(sum((arima_fcast_2[['mean']]-test)^2)/length(test))

#AICC
  arima_aicc_1[i] <-  arima_fit_1$aicc
  arima_aicc_2[i] <-  arima_fit_2$aicc

}

```

#### Mean RMSE value

```{r mean rmse,include=FALSE,eval=FALSE }
#RMSE Mean value of different models
# arima expanding 
mean(sqrt(colMeans(arima_mae_1^2,na.rm=TRUE))) 
# arima sliding 
mean(sqrt(colMeans(arima_mae_2^2,na.rm=TRUE)))
# ETS expanding 


#Calculate RMSE and plot per horizon
plot(1:31, sqrt(colMeans(arima_mae_1^2,na.rm=TRUE)), type="l",col="red",xlab="Horizon(Days)", ylab="RMSE",lty = 1,lwd = 3,main = "Root-square Forecast Error (RMSE) vs forecast horizon",ylim=c(10000,75000),cex.axis = 1.5,cex.lab = 1.5)
lines(1:31, sqrt(colMeans(arima_mae_2^2,na.rm=TRUE)), type="l",col="green",lty = 1,lwd = 3)
legend("topleft",legend=c("sARIMA - Expanding Window","sARIMA - Sliding Window" ),col=c("orange","blue"),lty=1,box.lty=0)


```
