---
title: "Web Traffic Time Series Forecasting -"
author: "Surendiran Rangaraj"
date: "11/7/2020"
output:
  html_document:
    fig_height: 5
    fig_width: 8
    toc: yes
    toc_depth: 1
    toc_float:
      collapsed: false
    number_sections: false
    theme: readable
    highlight: tango
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=FALSE}
library(fpp2) # time series analysis
library(tseries) # time series analysis
library(ggplot2) # visualization
library(forecast)  # time series analysis
library(TSA)  # time series analysis
library(dplyr) # data manipulation
library(readxl) # data input
library(xts) # time series analysis
library(tibble) # data wrangling
library(tidyr) # data wrangling
library(prophet) # time series analysis
library(zoo)
library(TSstudio)

```


```{r Read data}
#Read the Input data as Dataframe
webdata_path = 'C:/Users/rsure/Desktop/Uchicago/Time Series/Final Project/'
train_1_df = read.table(paste0(webdata_path,'train_1.csv'),sep=',',header = TRUE,skip =0,row.names = "Page")
#train_2_df = read.table(paste0(webdata_path,'train_2.csv'),sep=',',header = TRUE)
#key_1_df = read.csv2(paste0(webdata_path,'key_1.csv'),sep=',',header = TRUE)
#key_2_df = read.csv2(paste0(webdata_path,'key_2.csv'),sep=',',header = TRUE)

```

```{r data exploration}
#Basic Data
c(ncol(train_1_df),nrow(train_1_df))
head(train_1_df)
glimpse(train_1_df)


```

### Select Webpage for Analysis

Select the page India_en.wikipedia.org_all-access_all-agents,2016_FIFA_Club_World_Cup_en and  2016_Summer_Olympics_en for time series forecasting.

```{r select pages}
#
india_web_df = data.matrix(train_1_df["India_en.wikipedia.org_all-access_all-agents",])
fifa_web_df = data.matrix(train_1_df["2016_FIFA_Club_World_Cup_en.wikipedia.org_desktop_all-agents",])
olympics_web_df = data.matrix(train_1_df["2016_Summer_Olympics_en.wikipedia.org_desktop_all-agents",])

#Check for missing values 

print(paste0("Number of Missing Values of India webpage :",(sum(is.na(india_web_df))/(ncol(india_web_df)*nrow(india_web_df)))))
#india webpage
india_web_matrix = matrix(india_web_df, ncol = ncol(train_1_df), dimnames = NULL)
india_web_array=array(india_web_df)
length(india_web_array)
head(india_web_array)

#2016 FIFA World CUP
print(paste0("Number of Missing Values for 2016 FIFA:",(sum(is.na(fifa_web_df))/(ncol(fifa_web_df)*nrow(fifa_web_df)))))
fifa_web_matrix = matrix(fifa_web_df, ncol = ncol(train_1_df), dimnames = NULL)
fifa_web_array=array(fifa_web_df)
length(fifa_web_array)
head(fifa_web_array)

#2016 Summer Olympics

print(paste0("Number of Missing Values for 2016 Olympics:",(sum(is.na(olympics_web_df))/(ncol(olympics_web_df)*nrow(olympics_web_df)))))
olympics_web_matrix = matrix(olympics_web_df, ncol = ncol(train_1_df), dimnames = NULL)
olympics_web_array=array(olympics_web_df)
length(olympics_web_array)
head(olympics_web_array)

```

No Missing values found of the 550 Observations . 

### 2016 Summer Olympics Time series modeling 

```{R Plot Summer Olympics}

plot(olympics_web_array,type = 'l')

plot(olympics_web_array[1:100],type = 'l')

plot(olympics_web_array[1:14],type = 'l')

```

There is block intervention in our data. We will look at the time series data for 2016 Summer Olympics and perform basic modeling

### Convert to Time Series
First Let's convert it to time-series and then split it to train and test - we will train our models for the period 1st July 2015 till 30th Nov 2016 and we will forecast for 1st Dec 2016 to 31st Dec 2016

```{r convert ts}
# convert to time-series

#time_idx = seq(from = as.POSIXct("2015-07-01"),to = as.POSIXct("2016-12-31"), by = "day")
#olympics_web_ts = ts(olympics_web_array, start=c(2015,07),frequency = 365)

idx <- seq(as.Date("2015-07-01"), as.Date("2016-12-31"), by = "day")
olympics_web_ts = ts(olympics_web_array, frequency = 7)

head(olympics_web_ts)
tsdisplay(olympics_web_ts)

```

ACF and PACF clearly shows the data is not stationary 

```{r split train_test,include = FALSE}
# splitting to train and test
olympics_web_ts_train<-window(olympics_web_ts,start=c(2015,7), end=c(2016,5),frequency=365)
olympics_web_ts_test<-window(olympics_web_ts,start=c(2016,6),frequency=365)


#olympics_web_ts_train = olympics_web_ts['2015-07-01/2016-11-30']
#olympics_web_ts_test = olympics_web_ts['2016-12-01/2016-12-31']

# viewing the train and test
tail(olympics_web_ts_train) 
head(olympics_web_ts_test)

print(paste0("Total Observations in Training dataset : ", length(olympics_web_ts_train), " Total Observations in Test dataset : ", length(olympics_web_ts_test)))

autoplot(olympics_web_ts_train,xlab = "Time",ylab = "No of Visits", main = "Time Series plot of 2016 Olympics English Web page Visits")
autoplot(olympics_web_ts_test,xlab = "Time",ylab = "No of Visits", main = "Time Series plot of 2016 Olympics English Web page Visits")

```

```{r split}

split_ts <- ts_split(ts.obj = olympics_web_ts, sample.out = 31)

olympics_web_ts_train <- split_ts$train
olympics_web_ts_test <- split_ts$test

# viewing the train and test
tail(olympics_web_ts_train) 
head(olympics_web_ts_test)

print(paste0("Total Observations in Training dataset : ", length(olympics_web_ts_train), " Total Observations in Test dataset : ", length(olympics_web_ts_test)))

autoplot(olympics_web_ts_train,xlab = "Time",ylab = "No of Visits", main = "Time Series plot of 2016 Olympics English Web page Visits")
autoplot(olympics_web_ts_test,xlab = "Time",ylab = "No of Visits", main = "Time Series plot of 2016 Olympics English Web page Visits")

```
```{r stl}

##stl(olympics_web_ts)

olympics_web_ts %>% mstl() %>%
  autoplot() + xlab("Time")

msts(olympics_web_array,)

```


```{r test}
kpss.test(olympics_web_ts,null = 'Trend') # p-value = 0.01  < 0.05 # Reject Null Hypothesis => The process is not stationary 
adf.test(olympics_web_ts) # p-value = 0.07132 > 0.05  # Reject Null Hypothesis => The process is not stationary 

print(paste0("Boxcox lambda value for the entire dataset  is :" ,BoxCox.lambda(olympics_web_ts)))
print(paste0("Boxcox lambda value for the train dataset  is :" ,BoxCox.lambda(olympics_web_ts_train)))

```


```{r plotts data}

tsdisplay(olympics_web_ts_train,ylab="India.en Web page Visits",xlab="Time")
tsdisplay(olympics_web_ts_test,ylab="India.en Web page Visits",xlab="Time")

kpss.test(olympics_web_ts_train,null = 'Trend') # p-value = 0.01261  < 0.05 # Reject Null Hypothesis => The process is not stationary 
adf.test(olympics_web_ts_train) # p-value = 0.07705 > 0.05  # Accept Null Hypothesis => The process is not stationary 

print(paste0("Boxcox lambda value for the train dataset  is :" ,BoxCox.lambda(olympics_web_ts_train)))
      
```


<div class="alert alert-info">
 
 
</div>


**Kwiatkowski Phillips Schmidt Shin (KPSS) tests**

Null hypothesis: The process is stationary => Large p value

Alternative hypothesis: The process is nonstationary =>  Small p value

**Augmented Dickey Fuller test**

Null hypothesis: The process is nonstationary => Large p value

Alternative hypothesis: The process is stationary => Small p value

### Remove Outliers

```{r outlier, include= FALSE}
```

### Box-Cox Transformation

```{r box-cox}

lambda = BoxCox.lambda(olympics_web_ts)

olympics_web_ts_Boxcox = BoxCox(olympics_web_ts,lambda)

tsdisplay(olympics_web_ts_Boxcox)

plot(olympics_web_ts_Boxcox[1:30],type="l")

kpss.test(olympics_web_ts_Boxcox)

adf.test(olympics_web_ts_Boxcox)

```


###  Differencing

```{r nsdiff,include = TRUE}

# Apply Non-seasonal differencing
olympics_web_ts_Boxcox_sdiff <- diff(olympics_web_ts_Boxcox,7)
tsdisplay(olympics_web_ts_Boxcox_sdiff)
kpss.test(olympics_web_ts_Boxcox_sdiff)

# Apply Non-seasonal differencing
olympics_web_ts_Boxcox_diff <- diff(olympics_web_ts_Boxcox_sdiff)
# Apply Non-seasonal differencing
olympics_web_ts_Boxcox_diff2 <- diff(olympics_web_ts_Boxcox_diff)

# time series display of differenced data
tsdisplay(olympics_web_ts_Boxcox_diff)
tsdisplay(olympics_web_ts_Boxcox_diff2)

kpss.test(olympics_web_ts_Boxcox_diff) # Null hypothesis: The process is stationary => Large p value of 0.1
kpss.test(olympics_web_ts_Boxcox_diff2) # Null hypothesis: The process is stationary => Large p value of 0.1

```

### ARFIMA 

```{r arfima}

acf(olympics_web_ts,lag.max = 40)
acf(olympics_web_ts_Boxcox,lag.max = 150)

olympics_web_arfima = arfima(olympics_web_ts_Boxcox)
summary(olympics_web_arfima)

```


### ETS model

 ETS model and see how it turns out.

```{r ets}
olympics_web_ets_fit <- ets(olympics_web_ts,lambda = "auto")
olympics_web_ets_fit 

#accuracy(forecast(olympics_web_ets_fit,h=10),olympics_web_ts_test)

plot(forecast(olympics_web_ets_fit,h=10))

checkresiduals(olympics_web_ets_fit)

```


### Naive method

```{r naive}

checkresiduals(naive(olympics_web_ts))
mean(naive(olympics_web_ts)$residuals,na.rm = TRUE)

ts()

```

### Applying auto.arima

Applying the model suggested by auto.arima

```{r autoarima}

fr <- fourier(ts(olympics_web_array, frequency=365.25), K=5)

olympics_web_arima_fit <- auto.arima(olympics_web_ts,lambda = "auto",seasonal = TRUE,d=1,D=1)
olympics_web_arima_fit
```
```{r fitted}
plot(olympics_web_arima_fit$x, col="red", type="l")
lines(fitted(olympics_web_arima_fit), type="l",col="blue")

checkresiduals(olympics_web_arima_fit) # p-value = 0.001 < 0.05 ==> Reject Null hypothesis ==>  The TS is not independently distributed
```
```{r fitted}
plot(olympics_web_arima_fit$x[1:100], col="green", type="l")
lines(fitted(olympics_web_arima_fit)[1:100], type="l",col="red")

plot(olympics_web_arima_fit$x[400:450], col="green", type="l")
lines(fitted(olympics_web_arima_fit)[400:450], type="l",col="red")

plot(olympics_web_arima_fit$x[400:420], col="green", type="l")
lines(fitted(olympics_web_arima_fit)[400:420], type="l",col="red")

```

### auto arima with no seasonal differencing 

```{r autoarima}

fr <- fourier(ts(olympics_web_array, frequency=365.25), K=5)

olympics_web_arima_fit_ns <- auto.arima(olympics_web_ts,lambda = "auto",seasonal = TRUE,d=1,D=0)
olympics_web_arima_fit_ns
```

```{r merge,include = FALSE}
#merge(fitted(olympics_web_arima_fit),olympics_web_arima_fit$x, all = TRUE, by.x = "fitted(olympics_web_arima_fit)", by.y = "olympics_web_arima_fit$x")

cbind(fit=fitted(olympics_web_arima_fit),actual_data = olympics_web_arima_fit$x,diff=round((olympics_web_arima_fit$x - fitted(olympics_web_arima_fit)),2),diff_pct = ((olympics_web_arima_fit$x - fitted(olympics_web_arima_fit))/olympics_web_arima_fit$x)*100)


```
``` {r tbats}

tbats(olympics_web_ts,)

```

